---
title: "Billie Eilish: Chart Analysis"
format: html
editor: visual
---

**Billie Eilish** is an American singer, songwriter, and performer known for her unique sound, dark aesthetic, and emotionally raw lyrics. She rose to fame in 2015 at the age of 13 with her breakout song **“Ocean Eyes,”** which she released on SoundCloud and which quickly went viral. Her full name is **Billie Eilish Pirate Baird O’Connell**, and she was born on **December 18, 2001**, in Los Angeles, California.

# Billie Eilish: Song Rank Over Time

This plot visualizes the performance of Billie Eilish’s songs on the **Billboard Hot 100** over time.

```{r setup, include=FALSE}

library(tidyverse)
library(readr)
library(plotly)

charts <- read_csv("~/Desktop/data/charts.csv")

billie <- charts %>%
  filter(artist == "Billie Eilish")
```

```{r}
#| echo: false
ggplot(billie, aes(x = as.Date(date), y = rank, color = song)) +
  geom_line(linewidth = 0.5) +
  scale_y_reverse() +
  labs(
       x = "Date", y = "Chart Rank", color = "Song") 
  

```

# Billie Eilish: Heatmap of Songs Over Time

This heatmap visualizes the **Billboard chart activity** of Billie Eilish’s songs **week by week over time.**

```{r}
#| echo: false

ggplot(billie, aes(x = as.Date(date), y = song, fill = rank)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", direction = -1) +
  labs(
       x = "Date", y = "Song", fill = "Rank") +
  theme_minimal()


```

# Billie Eilish: Total Weeks on Chart by Song

This bar chart displays how many **total weeks** each Billie Eilish song spent on the **Billboard Hot 100** chart.

```{r}
#| echo: false
billie_weeks <- billie %>%
  group_by(song) %>%
  summarise(total_weeks = n()) %>%
  arrange(desc(total_weeks))

ggplot(billie_weeks, aes(x = reorder(song, total_weeks), y = total_weeks, fill = song)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  labs(
       x = "Song", y = "Weeks on Chart") +
  theme_minimal()

```

# Billie Eilish: Rank Distribution by Song

This chart visualizes the **distribution of chart rankings** for each Billie Eilish song on the **Billboard Hot 100** over its lifetime.

```{r}
#| echo: false
library(plotly)
p3 <- ggplot(billie, aes(x = song, y = rank, fill = song)) +
  geom_boxplot(show.legend = FALSE) +
  scale_y_reverse() +
  labs(
       x = "Song", y = "Rank") +
  theme(legend.position = "none",axis.text.x = element_text(angle = 75,hjust = 1))

ggplotly(p3)

```
