---
title: "Lady Gaga: Chart Analysis"
format: html
editor: visual
---

**Lady Gaga** (Stefani Germanotta) is an American singer, songwriter, and actress known for her bold fashion, powerful vocals, and hits like "Poker Face" and "Bad Romance". She gained fame with The Fame (2008) and has since released albums across pop, jazz, and electronic genres. Gaga is also an Oscar-winning actress, celebrated for her role in A Star Is Born (2018).

# Total Weeks on Chart per Song

This bar chart shows the **total number of weeks** each song by **Lady Gaga** spent on the **Billboard Hot 100**.

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)
library(plotly)
df <- read_csv("~/Desktop/data/charts.csv")
gaga_df <- df %>%
  filter(str_detect(artist, regex("Lady Gaga", ignore_case = TRUE))) %>%
  mutate(date = ymd(date), year = year(date))
```

```{r}
#| echo: false
g1 <- gaga_df %>%
  group_by(song) %>%
  summarise(max_weeks = max(`weeks-on-board`, na.rm = TRUE)) %>%
  arrange(desc(max_weeks)) %>%
  ggplot(aes(x = reorder(song, max_weeks), y = max_weeks)) +
  geom_col(fill = "mediumpurple") +
  coord_flip() +
  labs( x = "Song", y = "Weeks")

g2 <- gaga_df %>%
  group_by(song) %>%
  summarise(best_peak = min(`peak-rank`, na.rm = TRUE)) %>%
  arrange(best_peak) %>%
  ggplot(aes(x = reorder(song, best_peak), y = best_peak)) +
  geom_col(fill = "salmon") +
  coord_flip() +
  labs( x = "Song", y = "Peak Rank")

g3 <- gaga_df %>%
  count(year) %>%
  ggplot(aes(x = factor(year), y = n)) +
  geom_col(fill = "skyblue") +
  labs(title = "", x = "Year", y = "Entries")



g1

```

# Best Peak Rank per Song

This chart displays the **highest (peak) Billboard Hot 100 position** that each of Lady Gaga's songs achieved.

```{r}
#| echo: false
g2

```

# Number of Chart Entries per Year

This bar chart displays the **number of unique songs by Lady Gaga** that entered the **Billboard Hot 100 chart** each year.

```{r}
#| echo: false
ggplotly(g3)
```
