---
title: "The Weeknd: Chart Performance Analysis"
format: html
editor: visual
---

**The Weeknd** (Abel Tesfaye) is a Canadian singer, songwriter, and producer known for his distinctive voice and genre-blending music. He gained fame with his 2011 mixtapes *House of Balloons* and later global hits like *"Can't Feel My Face"*, *"Blinding Lights"*, and *"Starboy"*. His music blends R&B, pop, and synthwave, and he's known for his enigmatic persona and immersive performances.

# The Weeknd: Song Rank Over Time

This plot visualizes the performance of **The Weeknd’s songs** on the **Billboard Hot 100** over time.

```{r setup, include=FALSE}
library(tidyverse)
library(readr)
library(plotly)


charts <- read_csv("~/Desktop/data/charts.csv")
weeknd <- charts %>%
  filter(artist == "The Weeknd")
```

```{r}
#| echo: false

p1 <- ggplot(weeknd, aes(x = as.Date(date), y = rank, color = song)) +
  geom_line(linewidth = 0.5) +
  scale_y_reverse() +
  labs(
       x = "Date", y = "Chart Rank", color = "Song") +
  theme_minimal()

ggplotly(p1)


```

# The Weeknd: Rank Distribution by Song

This plot visualizes the **distribution of chart rankings** for each song by **The Weeknd** on the **Billboard Hot 100** throughout its time on the chart.

```{r}
#| echo: false
ggplot(weeknd, aes(x = rank, y = song, fill = song)) +
  geom_boxplot(show.legend = FALSE) +
  labs(
       y = "Song", x = "Chart Rank") +
  theme_minimal()

```
